---
title: "Confidence Intervals and NHST"
author: "[your name]"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("infer")     #tools to easily code confidence intervals and NHST
library("tidyverse") #data wrangling tools such as filter
```

The materials for this homework assignment come from Chapters 8 and 9 of **Statistical Inference via Data Science:** *A Modern Dive into R and the Tidyverse* along with the class lectures.

* [https://moderndive.com/8-confidence-intervals.html](https://moderndive.com/8-confidence-intervals.html)
* [https://moderndive.com/9-hypothesis-testing.html](https://moderndive.com/9-hypothesis-testing.html)

# Data

```{r, message = FALSE, warning = FALSE}
df <- read_csv("demographics_survey_data.csv")
```

# Confidence Intervals

## Friendships

The survey question was "How many friends do you talk to at least once per month?"

(A) View the `summary` of the `friends` variable from the `df` data frame.

```{r}

```

(B) Create a *bootstrap distribution* of the `friends` variable.

```{r}

```

(c) Plot the bootstrap distribution `visualize` (from the `infer` package) along with the confidence interval layer for a 95% confidence interval.

```{r}

```

(D)  Finally, compute the endpoints of the 95% confidence interval, and type a complete sentence to describe the results.

```{r}

```



## Recommend UCM?

The survey question was "On a scale from 0 ("would not") to 100 ("highly"), how much would you recommend UC Merced to people that are about to apply to college?"

(A) View the `summary` of the `recommendUCM` variable from the `df` data frame.

```{r}

```

(B) Create a *bootstrap distribution* of the `friends` variable.

```{r}

```

(c) Plot the bootstrap distribution `visualize` (from the `infer` package) along with the confidence interval layer for a 95% confidence interval.

```{r}

```

(D)  Finally, compute the endpoints of the 95% confidence interval, and type a complete sentence to describe the results.

```{r}

```



# Hypothesis Testing

## Procrastination vs Partying

The survey questions were

* "Do you consider yourself to be a procrastinator?"
* "Do you plan to attend at least one party per month this semester?"

* colloquial hypothesis: Party goers are more likely to be procrastinators
* null hypothesis: Procrastinator proportion is the same regardless of party going plans
* alternative hypothesis: Party goers are more likely to be procrastinators

* predictor variable (categorical): `parties`
* response variable (categorical): `procrastinator`

In this scenario, our one-sided hypothesis test looks at a difference of two proportions.

$$\begin{array}{rrcl}
  H_{0}: p_{Y} - p_{N} & = & 0 \\
  H_{a}: p_{Y} - p_{N} & > & 0 \\
\end{array}$$

(a) Compute the observed difference in proportions for the procrastination grouped by parties

```{r}
# table(demo_df$procrastinator)
# table(demo_df$parties)
```

(b) Build a null distribution toward the hypothesis test.

```{r}

```

(c) Use the `visualize` function to graph the null distribution and shade the p-value.

```{r}

```

(d) Compute the p-value for the hypothesis test, and then write a sentence to describe the conclusion of the hypothesis test.

```{r}

```



## Homesickness vs UC Merced as a top-choice school

The survey questions were

* "On a scale from 0 to 100, how homesick are you?"
* "When applying and accepting college, was UC Merced your top choice?"

* null hypothesis: Homesickness levels are the same whether or not UC Merced was the student's top choice.
* alternative hypothesis: Homesickness levels are different whether or not UC Merced was the student's top choice.

* predictor variable (categorical): `UCMtopChoice`
* response variable (numerical): `homesickness`

In this scenario, our one-sided hypothesis test looks at a difference of two means.

$$\begin{array}{rrcl}
  H_{0}: \mu_{Y} - \mu_{N} & = & 0 \\
  H_{a}: \mu_{Y} - \mu_{N} & \neq & 0 \\
\end{array}$$

(a) Compute the observed difference in means for the homesickness grouped by students who felt UC Merced was their top choice.

```{r}
# table(demo_df$UCMtopChoice)
# summary(demo_df$homesickness)
```

(b) Build a null distribution toward the hypothesis test.

```{r}

```

(c) Use the `visualize` function to graph the null distribution and shade the p-value.

```{r}

```

(d) Compute the p-value for the hypothesis test, and then write a sentence to describe the conclusion of the hypothesis test.

```{r}

```


# Wrap Up

When you are done and your code is free of errors, use the `Knit` button to `Knit to HTML`.  Upload the HTML file back to CatCourses.